Bayesian Sparsification of $\cplx$-valued networks

### 2019-11-11 paper/paper.tex
finished this editing iteration

Radio Frequency power amplifiers, being physical electronic devices, have imperfections such
as parasitic inductance and capacitance. This coupled with them being operated in saturation,
make their gain respond non-linearly to the input signal, adding undesirable interference and
spectral spreading. Although these effects can be minimized by careful circuit design, they
cannot be eliminated entirely \cite{citation_needed}. Digital signal predistortion aims at
linearizing the amplifier, by introducing signal-dependent compensating component to the input.
At the same time the non-linearities should be corrected on-device efficiently and with low
latency in order to reduce cross-channel interference and enable higher bandwidth utilization
in wireless transmitters.

@@

Bayesian inference is a very rich field of research with a large body of knowledge and techniques
that enable practical quantification of predictive, parameter, or model uncertainty and its
propagation into solutions of the problems of interest. It is not safeguarded against overfitting
and modelling biases: there are quite a handful of assumptions made every time a Bayesian model
is setup for inference. We need to have a prior belief, the model and its interaction with
the data (the likelihood), and even the posterior: although the Bayes formula is explicit in
how the posterior is derived, in all problems of at least some practical significance the
exact posterior inference involves analytically intractable expectations, compelling the use
of approximations. In addition to this, even if the likelihood and prior come from conjugate
families, making the posterior explicit, the questions of interest are usually concerned with
inference using the posterior predictive distribution, which is factorized as the conditional
predictive distribution given the model and new observation, and the posterior of the model
after observing the data.

% criticism and discussion
Despite this, $\cplx$-networks remain much less popular than $\real$-networks for several
reasons. One possible reason is that a $\cplx$-layer is identical to a paired $\real$-layer,
based on the algebraic equivalence between $2 \times 2$ real matrices and complex numbers.
Due to this constraint $\cplx$-networks have less expressive power for the same computation
complexity compared to a $\real$-layer with twice as many inputs and outputs.
%
Another one is that a $\cplx$-network is not a true complex-valued function of a complex
argument: common non-linearities are pointwise applications of $\real$-valued activations
\cite{trabelsi_deep_2017}.
%
Third conceptual issue is that a real-valued loss as a $\cplx\to\real$ function fails
Cauchy-Riemann conditions, and so is not $\cplx$-differentiable. Since the network has to
be trained by a gradient-based method (stochastic), this issue is overcome by $\cplx\real$,
or Wirtinger calculus, which treats the complex argument $z$ and its conjugate $\conj{z}$
as explicit independent variables. For non-holomorphic functions this approach makes taking
their derivatives essentially indistinguishable from calculus on $\real^2$.
%
The last reason is the lack of native support by the major deep learning and auto-differentiation
frameworks.

Practical performance of variational dropout techniques was studied in \cite{citation_needed}
and \cite{citation_needed}. The most fierce critique of Bayesian approach towards model
sparsification was voiced in \cite{gale_state_2019}, where it was show that it yielded
inconsistent results in text translation tasks and systematically underperformed simpler
techniques such as norm pruning and $\ell_0$-regularization, \cite{louizos_learning_2017}.

The latter approach introduces stochastic masks one the parameters, which have hard-concrete
mask distribution. It is closely related to gumbel-softmax reparameterization trick,
\cite{kingma_auto-encoding_2014,jang_categorical_2017}: the mask values are, essentially,
$[0,1]$ clamped scaled sigmoids, that are randomly shifted relative to a learnable location
parameter, that indicates the degree of weight's relevance. The penalty in the objective is
the expected number of nonzero parameters, and is a smooth approximation to $\ell_0$ norm.
This approach is related to stochastic, rather than bayesian networks: it is formulated in
terms of empirical risk minimization of the negative $\log$-likelihood term in \eqref{eq:elbo}.

@@

In this section we derive the exact gradient of the KL-divergence term \eqref{eq:improper-kl-div-real}
for $\mathbb{R}$-dropout in case of a log-uniform prior. Even though the final expression for
the penalty term itself involves a derivative of a hypergeometric function, and the gradient,
essentially, relies on computing the error function, the following analysis provides independent
verification of the correctness of the approximation, proposed in \cite{molchanov_variational_2017}.
The logic of the derivation follows \cite{lapidoth_capacity_2003}.
% \cite{pav_moments_2015} correctly notice that \cite[p. 2466]{lapidoth_capacity_2003} has a missing $\log{2}$ term.

Let $(z_i)_{i=1}^m \sim \mathcal{N}(0, 1)$ iid and $
  (\mu_i)_{i=1}^m \in \mathbb{R}
$. The random variable $
  W = \sum_i (\mu_i + z_i)^2
$ is said to be $\chi^2_m(\lambda)$, i.e. noncentral $\chi^2$ with shape $m$ and noncentrality
parameter $\lambda = \sum_i \mu_i^2$. Its density is given by, \cite{citation_needed}
$$
f_W(x)
  % = e^{- \tfrac\lambda2} \sum_{n \geq 0} \tfrac{\bigl(\tfrac\lambda2\bigr)^n}{n!}
  %   \tfrac{
  %     x^{n + \tfrac{m}2 - 1} e^{-\tfrac{x}2}
  %   }{
  %     2^{n + \tfrac{m}2} \Gamma(n + \tfrac{m}2)
  %   }
  % = e^{- \tfrac\lambda2} \sum_{n \geq 0} \tfrac{\bigl(\tfrac\lambda2\bigr)^n}{n!}
  %     \tfrac{
  %         \bigl(\tfrac{x}2\bigr)^{n + \tfrac{m}2 - 1} e^{-\tfrac{x}2}
  %     }{
  %         2 \Gamma(n + \tfrac{m}2)
  %     }
  = \frac12 e^{- \tfrac{x + \lambda}2} \bigl(\tfrac{x}2\bigr)^{\tfrac{m}2 - 1}
    \sum_{n \geq 0} \tfrac{
      \bigl(\tfrac{x \lambda}4\bigr)^n
    }{
      n! \Gamma(n + \tfrac{m}2)
    }
  % = \frac12 e^{- \tfrac{x + \lambda}2}
  % \bigl(\tfrac{x}\lambda\bigr)^{\tfrac{m}4 - \tfrac12}
  % \bigl(\tfrac{x \lambda}4\bigr)^{\tfrac{m}4 - \tfrac12}
  % \sum_{n \geq 0} \tfrac{
  %         \bigl(\tfrac{x \lambda}4\bigr)^n
  %     }{
  %         n! \Gamma(n + \tfrac{m}2)
  %     }
  % = \frac12 e^{- \tfrac{x + \lambda}2}
  % \bigl(\tfrac{x}\lambda\bigr)^{\tfrac{m}4 - \tfrac12}
  % \bigl(\tfrac{\sqrt{x \lambda}}2\bigr)^{\tfrac{m}2 - 1}
  % \sum_{n \geq 0} \tfrac{
  %         \bigl(\tfrac{x \lambda}4\bigr)^n
  %     }{
  %         n! \Gamma(n + \tfrac{m}2)
  %     }
  % = \frac12 e^{- \tfrac{x + \lambda}2}
  % \bigl(\tfrac{x}\lambda\bigr)^{\tfrac{m - 2}4}
  % I_{\bigl(\tfrac{m}2 - 1\bigr)}(\sqrt{x \lambda})
  \,. $$
% where $I_k$ is the modified Bessel function of the first kind
% $$
% I_k(s)
%   = \Bigl(\frac{s}2\Bigr)^k
%       \sum_{n \geq 0} \tfrac{
%           \bigl( \tfrac{s}2 \bigr)^{2n}
%       }{n! \Gamma(n + k + 1)}
%   \,. $$

In the real multiplicative noise case we need to compute the kl-divergence penalty, the key
term of which is the expected logarithm of $W$:
\begin{align}
  \mathbb{E}_{W\sim \chi^2_m(\lambda)} \log W
    &= \int_0^\infty f_W(x) \log x dx
    \notag \\
    &= \frac12 e^{- \tfrac\lambda2}
    \int_0^\infty \sum_{n \geq 0} \biggl(
        \tfrac{\bigl(\tfrac\lambda2\bigr)^n}{n!}
        \tfrac{e^{- \tfrac{x}2}}{\Gamma(n + \tfrac{m}2)}
        \bigl(\tfrac{x}2\bigr)^{n + \tfrac{m}2 - 1}
    \biggr) \log x dx
    % = [\text{ absolute summability and Fubini, or any other conv. thm for integrals of non-negative integrals}]
    % = e^{- \tfrac\lambda2} \sum_{n \geq 0}
    %     \tfrac{\bigl(\tfrac\lambda2\bigr)^n}{n!}
    %     \tfrac1{\Gamma(n + \tfrac{m}2)}
    % \int_0^\infty
    %     e^{- \tfrac{x}2} \bigl(\tfrac{x}2\bigr)^{n + \tfrac{m}2 - 1}
    %     (\log2 + \log \tfrac{x}2) \tfrac{dx}2
    % = [u = \tfrac{x}2]
    \notag \\
    &= e^{- \tfrac\lambda2} \sum_{n \geq 0}
        \tfrac{\bigl(\tfrac\lambda2\bigr)^n}{n!}
    \tfrac1{\Gamma(n + \tfrac{m}2)}
    \int_0^\infty
        e^{- u} u^{n + \tfrac{m}2 - 1}
        (\log2 + \log u) du
    % = [\text{definitions:} \Gamma, \psi]
    % = e^{- \tfrac\lambda2} \sum_{n \geq 0}
    %     \tfrac{\bigl(\tfrac\lambda2\bigr)^n}{n!}
    % (\psi(n + \tfrac{m}2) + \log2)
    \notag \\
    &= \log{2}
    + e^{- \tfrac\lambda2} \sum_{n \geq 0}
        \tfrac{\bigl(\tfrac\lambda2\bigr)^n}{n!}
        \psi(n + \tfrac{m}2)
    \,,
\end{align}
where $\psi(x)$ is the digamma function, i.e.
$$
\psi(x)
  = \frac{d}{dx} \log \Gamma(x)
  = \frac1{\Gamma(x)}
    \int_0^\infty
      u^{x-1} e^{-u} \log{u}
    \, du
  \,. $$
The following properties of the digamma function will be used in the our analysis:
$
  \psi(z+1) = \psi(z) + \tfrac1z
$ and $
  \psi(\tfrac12) = -\gamma - 2\log 2
$. If we put
$$
g_m(x)
  % alpha: e^{-x} \sum_{n=0}^{\infty} \frac{x^n}{n!} \psi(n + m / 2)
  % = e^{-x} \sum_{n \geq 0} \frac{x^n}{n! \Gamma(n + \tfrac{m}2)}
  %     \int_0^\infty e^{-t} t^{n + \tfrac{m}2-1} \log{t} dt
  = e^{-x} \sum_{n \geq 0} \frac{x^n}{n!} \psi(n + \tfrac{m}2)
  \,, $$
then the desired expectation $
  \mathbb{E}_{z \sim \mathcal{N}(0, 1)}
    \log{\bigl\lvert z + \tfrac1{\sqrt{\alpha}} \bigr\rvert^2}
$ is equal to $
  \log{2} + g_1(\tfrac1{2\alpha})
$.

We can differentiate the series in $g_m(x)$, since the sum converges everywhere
on $\mathbb{R}$. Indeed, it is a power series featuring nonnegative coefficients,
which is dominated by $
  \sum_{n \geq 1} \tfrac{x^n}{n!} \log{(n+\tfrac{m}2)}
$, because $\psi(x)\leq \log x - \tfrac1{2x}$. By the ratio test, the dominating
series has infinite radius:
$$
\lim_{n\to\infty}
  \biggl\lvert
    \frac{
      n! x^{n+1} \log{(n + 1 + \tfrac{m}2)}
    }{
      x^n \log{(n + \tfrac{m}2)} (n+1)!
    }
  \biggr\rvert
  = \lim_{n\to\infty}
    \lvert x \rvert 
    \biggl\lvert
      \frac{
        \log{(n + 1 + \tfrac{m}2)}
      }{
        \log{(n + \tfrac{m}2)} (n+1)
      }
    \biggr\rvert
  % = \lim_{n\to\infty}
  %   \lvert x \rvert 
  %   \biggl\lvert
  %     \frac{
  %       n + \tfrac{m}2
  %     }{
  %       (n + 1 + \tfrac{m}2)((n+1) + (n + \tfrac{m}2) \log{(n + \tfrac{m}2)})
  %     }
  %   \biggr\rvert
  = 0 < 1
  \,, $$
since
$$
\frac{\log{(x + a + 1)}}{x \log{(x + a)}}
  \sim \frac{\log{(x+1)}}{(x-a) \log{x}}
  \sim \frac{\log{(x+1)}}{x \log{x}}
  \sim \frac{\tfrac1{(x+1)}}{1 + \log{x}}
  \to 0
  \,. $$
A theorem from calculus states, that the formal series derivative (integral)
coincides with the derivative (integral) of the function, corresponding to
the power series (everywhere on the convergence region). And the convergence
regions of derivative (integral) coincide with the region of the original
power series.

Differentiating the series within $g_m$ yields
$$
\frac{d}{d x} g_m(x)
  % = e^{-x} \sum_{n\geq 1} \frac{x^{n-1}}{(n-1)!} \psi(n + \tfrac{m}2) - g_m(x)
  % = e^{-x} \sum_{n\geq 0} \frac{x^n}{n!} \psi(n + \tfrac{m}2 + 1) - g_m(x)
  % = e^{-x} \sum_{n\geq 0} \frac{x^n}{n!} (
  %     \psi(n + \tfrac{m}2 + 1) - \psi(n + \tfrac{m}2)
  % )
  % = e^{-x} \sum_{n\geq 0} \frac{x^n}{n!} \tfrac1{n + \tfrac{m}2}
  = e^{-x} \sum_{n\geq 0} \frac{x^n}{n!} (n + \tfrac{m}2)^{-1}
  = \cdots
  \,. $$
Now, observe that $
    e^t  = \sum_{n\geq 0} \tfrac{t^n}{n!}
$ on $\mathbb{R}$, for $\alpha\neq 0$ we have $
  \int_0^x t^{\alpha-1} dt
    = \tfrac{x^\alpha}{\alpha}
$ and that
$$
\sum_{n\geq 0} \int_0^x \frac{t^{n+\alpha-1}}{n!} dt
  = \int_0^x \sum_{n\geq 0} \frac{t^{n+\alpha-1}}{n!} dt
  = \int_0^x t^{\alpha-1} e^t dt
  \,, $$
by the monotone convergence theorem on $
  ([0, x], \mathcal{B}([0, x]), dx)
$ whence
$$
\cdots
  = x^{-\tfrac{m}2} e^{-x} \sum_{n\geq 0}
      \frac{x^{n + \tfrac{m}2}}{n!} (n + \tfrac{m}2)^{-1}
  % = x^{-\tfrac{m}2} e^{-x} \sum_{n\geq 0}
  %     \frac1{n!} \int_0^x t^{n + \tfrac{m}2 - 1} dt
  % = x^{-\tfrac{m}2} e^{-x}
  %     \int_0^x t^{\tfrac{m}2 - 1} \sum_{n\geq 0} \frac{t^n}{n!} dt
  = x^{-\tfrac{m}2} e^{-x}
      \int_0^x t^{\tfrac{m}2 - 1} e^t dt
  = \cdots
  \,. $$
The substitution $u^2 = t$ on $[0, \infty]$ with $2u du = dt$ results in
$$
\int_0^x t^{\tfrac{m}2 - 1} e^t  dt
  % = \int_0^{\sqrt{x}} u^{m - 2} e^{u^2} 2 u du
  = 2 \int_0^{\sqrt{x}} u^{m - 1} e^{u^2} du
  \,, $$
whence
$$
\frac{d}{d x} g_m(x)
  % = x^{-\tfrac{m}2} e^{-x}
  %     \int_0^x t^{\tfrac{m}2 - 1} e^t dt
  = 2 x^{-\tfrac{m}2} e^{-x}
      \int_0^{\sqrt{x}} u^{m - 1} e^{u^2} du
  \,. $$
In particular, for $m=1$ we get
$$
g_1(x)
  = - \gamma - 2 \log{2}
  + e^{-x} \sum_{n\geq 0} \frac{x^n}{n!} \sum_{p=1}^n \frac2{2p - 1}
  \,, $$
and the derivative of the KL-term can be computed thus 
$$
\frac{d}{d x} g_1(x)
  = 2 \tfrac{F(\sqrt{x})}{\sqrt{x}}
  \,, $$
using Dawson's integral $
  F \colon \mathbb{R} \to \mathbb{R}
  \colon x \mapsto e^{-x^2} \int_0^x e^{u^2} du
$ -- a special function, implemented in most scientific computing libraries, e.g.
``scipy.special.dawsn''.

The derivative wrt $\alpha$ is 
$$
\tfrac{d}{d \alpha} g_1(\tfrac1{2\alpha})
  = -\tfrac1{2 \alpha^2} g_1'(\tfrac1{2\alpha})
  % = -\tfrac1{2 \alpha^2} 2 \tfrac{F(\sqrt{\tfrac1{2\alpha}})}{\sqrt{\tfrac1{2\alpha}}}
  % = -\tfrac1{2 \alpha^2} 2 \tfrac{F(\tfrac1{\sqrt{2\alpha}})}{\tfrac1{\sqrt{2\alpha}}}
  = -\tfrac1{\alpha} \sqrt{\tfrac2{\alpha}} F(\tfrac1{\sqrt{2\alpha}})
  \,. $$
However, since $\alpha$ is nonnegative, it is typically parameterized via its logarithm
and computed when needed. Thus, in particular, the gradient of the divergence penalty w.r.t
$\log \alpha$ is

\begin{equation}  \label{eq:real-kl-div-deriv-log}
  \frac{d}{d\log \alpha}
    \tfrac12 \mathbb{E}_{z \sim \mathcal{N}(0,1)}
    \log \lvert z + \tfrac1{\sqrt{\alpha}} \rvert^2
    = \frac12 \frac{d\alpha}{d\log \alpha}
      \frac{d}{d\alpha} \bigl(\mathbb{E}\cdots \bigr)
    % = - \frac\alpha2 \tfrac1{\alpha}
    %   \sqrt{\tfrac2{\alpha}} F(\tfrac1{\sqrt{2\alpha}})
    = - \tfrac1{\sqrt{2\alpha}} F(\tfrac1{\sqrt{2\alpha}})
    \,.
\end{equation}

>>>>

Radio Frequency Power Amplifiers (PA) are devices used to increase the power of a signal for
latter transmission via amplitude modulation over radio. One of the most important problems
in wireless transmission is distortion caused by imperfections the amplifier, which degrades
the throughput. Although careful circuit design can mitigate undesirable effects, they cannot
be eliminated entirely \cite{citation_needed}. \textit{Digital Signal Predistortion} (DPD)
aims at linearizing the amplifier by introducing computed on-the-fly compensation to the
input that pushes the signal into the linear region of the PA. Practically, the method helps
reduce cross-channel interference and enables higher bandwidth utilization in wireless
transmitters, \cite{citation_needed}. The main challenge of DPD methods is efficient
compensation on-device (with low-latency).

@@

Bayesian inference is a very rich field of research with a large body of knowledge and techniques
that enable practical quantification of predictive, parameter, or model uncertainty and its
propagation into solutions of the problems of interest.

% criticism and discussion
based on the algebraic equivalence between $2 \times 2$ real matrices and complex numbers.
Due to this constraint $\cplx$-networks have less expressive power for the same computation
complexity compared to a $\real$-layer with twice as many inputs and outputs.
%
The most fierce critique of Bayesian approach towards model
sparsification was voiced in \cite{gale_state_2019}, where it was show that it yielded
inconsistent results in text translation tasks and systematically underperformed simpler
techniques such as norm pruning and $\ell_0$-regularization, \cite{louizos_learning_2017}.

The latter approach introduces stochastic masks one the parameters, which have hard-concrete
mask distribution. It is closely related to gumbel-softmax reparameterization trick,
\cite{kingma_auto-encoding_2014,jang_categorical_2017}: the mask is a stretched sigmoid,
clamped to $[0,1]$, that takes random input shifted relative to a learnable location
parameter, which indicates the degree of relevance. The penalty in the objective is a
smooth approximation to $\ell_0$ norm, i.e. the expected number of nonzero parameters.

@@

In this section we derive the exact gradient of \eqref{eq:improper-kl-div-real}. Even
though, ultimately, the divergence involves a hypergeometric function \todo{which one},
and its gradient requires evaluating the Gauss error function, the analysis provides
independent verification of the approximation, proposed in \cite{molchanov_variational_2017}.
The logic of the derivation follows \cite{lapidoth_capacity_2003}.
% \cite{pav_moments_2015} correctly notice that \cite[p. 2466]{lapidoth_capacity_2003} has a missing $\log{2}$ term.

For $(z_i)_{i=1}^m \sim \mathcal{N}(0, 1)$ iid and $
  (\mu_i)_{i=1}^m \in \mathbb{R}
$. The random variable $W = \sum_i (\mu_i + z_i)^2$ noncentral $\chi^2$ with shape $m$
and noncentrality parameter $\lambda = \sum_i \mu_i^2$, i.e. $W\sim \chi^2_m(\lambda)$.
Its density is
\begin{equation}. \label{eq:nonchisq-density}
  f_W(x)
    % = e^{- \tfrac\lambda2} \sum_{n \geq 0} \tfrac{\bigl(\tfrac\lambda2\bigr)^n}{n!}
    %   \tfrac{
    %     x^{n + \tfrac{m}2 - 1} e^{-\tfrac{x}2}
    %   }{
    %     2^{n + \tfrac{m}2} \Gamma(n + \tfrac{m}2)
    %   }
    % = e^{- \tfrac\lambda2} \sum_{n \geq 0} \tfrac{\bigl(\tfrac\lambda2\bigr)^n}{n!}
    %     \tfrac{
    %         \bigl(\tfrac{x}2\bigr)^{n + \tfrac{m}2 - 1} e^{-\tfrac{x}2}
    %     }{
    %         2 \Gamma(n + \tfrac{m}2)
    %     }
    = \frac12 e^{- \tfrac{x + \lambda}2} \bigl(\tfrac{x}2\bigr)^{\tfrac{m}2 - 1}
      \sum_{n \geq 0} \tfrac{
        \bigl(\tfrac{x \lambda}4\bigr)^n
      }{
        n! \Gamma(n + \tfrac{m}2)
      }
    % = \frac12 e^{- \tfrac{x + \lambda}2}
    % \bigl(\tfrac{x}\lambda\bigr)^{\tfrac{m}4 - \tfrac12}
    % \bigl(\tfrac{x \lambda}4\bigr)^{\tfrac{m}4 - \tfrac12}
    % \sum_{n \geq 0} \tfrac{
    %         \bigl(\tfrac{x \lambda}4\bigr)^n
    %     }{
    %         n! \Gamma(n + \tfrac{m}2)
    %     }
    % = \frac12 e^{- \tfrac{x + \lambda}2}
    % \bigl(\tfrac{x}\lambda\bigr)^{\tfrac{m}4 - \tfrac12}
    % \bigl(\tfrac{\sqrt{x \lambda}}2\bigr)^{\tfrac{m}2 - 1}
    % \sum_{n \geq 0} \tfrac{
    %         \bigl(\tfrac{x \lambda}4\bigr)^n
    %     }{
    %         n! \Gamma(n + \tfrac{m}2)
    %     }
    % = \frac12 e^{- \tfrac{x + \lambda}2}
    % \bigl(\tfrac{x}\lambda\bigr)^{\tfrac{m - 2}4}
    % I_{\bigl(\tfrac{m}2 - 1\bigr)}(\sqrt{x \lambda})
% where $I_k$ is the modified Bessel function of the first kind
% I_k(s)
%   = \Bigl(\frac{s}2\Bigr)^k
%       \sum_{n \geq 0} \tfrac{
%           \bigl( \tfrac{s}2 \bigr)^{2n}
%       }{n! \Gamma(n + k + 1)}
In \eqref{eq:ard-kl-div-real} we need the expected logarithm of $W$:
\begin{align}
  \mathbb{E}_{W\sim \chi^2_m(\lambda)} \log W
    &= \int_0^\infty f_W(x) \log x dx
    &= \frac12 e^{- \tfrac\lambda2}
    \int_0^\infty \sum_{n \geq 0} \biggl(
        \tfrac{\bigl(\tfrac\lambda2\bigr)^n}{n!}
        \tfrac{e^{- \tfrac{x}2}}{\Gamma(n + \tfrac{m}2)}
        \bigl(\tfrac{x}2\bigr)^{n + \tfrac{m}2 - 1}
    \biggr) \log x dx
    % = [\text{ absolute summability and Fubini, or any other conv. thm for integrals of non-negative integrals}]
    % = e^{- \tfrac\lambda2} \sum_{n \geq 0}
    %     \tfrac{\bigl(\tfrac\lambda2\bigr)^n}{n!}
    %     \tfrac1{\Gamma(n + \tfrac{m}2)}
    % \int_0^\infty
    %     e^{- \tfrac{x}2} \bigl(\tfrac{x}2\bigr)^{n + \tfrac{m}2 - 1}
    %     (\log2 + \log \tfrac{x}2) \tfrac{dx}2
    % = [u = \tfrac{x}2]
    &= e^{- \tfrac\lambda2} \sum_{n \geq 0}
        \tfrac{\bigl(\tfrac\lambda2\bigr)^n}{n!}
    \tfrac1{\Gamma(n + \tfrac{m}2)}
    \int_0^\infty
        e^{- u} u^{n + \tfrac{m}2 - 1}
        (\log2 + \log u) du
    % = [\text{definitions:} \Gamma, \psi]
    % = e^{- \tfrac\lambda2} \sum_{n \geq 0}
    %     \tfrac{\bigl(\tfrac\lambda2\bigr)^n}{n!}
    % (\psi(n + \tfrac{m}2) + \log2)
    &= \log{2}
    + e^{- \tfrac\lambda2} \sum_{n \geq 0}
        \tfrac{\bigl(\tfrac\lambda2\bigr)^n}{n!}
        \psi(n + \tfrac{m}2)
where $\psi(x)$ is the digamma function, i.e.
\begin{equation}  \label{eq:digamma}
  \psi(x)
    = \frac{d}{dx} \log \Gamma(x)
    = \frac1{\Gamma(x)}
      \int_0^\infty
        u^{x-1} e^{-u} \log{u}
      \, du
Note that $\psi$ satisfies: $
  \psi(z+1) = \psi(z) + \tfrac1z
  \psi(x)\leq \log x - \tfrac1{2x}
$, and $
  \psi(\tfrac12) = -\gamma - 2\log 2
$ ($\gamma$ is Euler's constant). If we put
\begin{equation}  \label{eq:g-m-series}
  g_m(x)
    % alpha: e^{-x} \sum_{n=0}^{\infty} \frac{x^n}{n!} \psi(n + m / 2)
    % = e^{-x} \sum_{n \geq 0} \frac{x^n}{n! \Gamma(n + \tfrac{m}2)}
    %     \int_0^\infty e^{-t} t^{n + \tfrac{m}2-1} \log{t} dt
    = e^{-x} \sum_{n \geq 0} \frac{x^n}{n!} \psi(n + \tfrac{m}2)
then the desired expectation in \eqref{eq:ard-kl-div-real} equals $
  \log{2} + g_1(\tfrac1{2\alpha})
$.

% We can differentiate the series in $g_m(x)$, since the sum converges on $\mathbb{R}$.
% Indeed, it is a power series featuring nonnegative coefficients, which is dominated by $
%   \sum_{n \geq 1} \tfrac{x^n}{n!} \log{(n+\tfrac{m}2)}
% $. By the ratio test, the dominating
% series has infinite radius:
% $$
% \lim_{n\to\infty}
%   \biggl\lvert
%     \frac{
%       n! x^{n+1} \log{(n + 1 + \tfrac{m}2)}
%     }{
%       x^n \log{(n + \tfrac{m}2)} (n+1)!
%     }
%   \biggr\rvert
%   = \lim_{n\to\infty}
%     \lvert x \rvert 
%     \biggl\lvert
%       \frac{
%         \log{(n + 1 + \tfrac{m}2)}
%       }{
%         \log{(n + \tfrac{m}2)} (n+1)
%       }
%     \biggr\rvert
%   % = \lim_{n\to\infty}
%   %   \lvert x \rvert 
%   %   \biggl\lvert
%   %     \frac{
%   %       n + \tfrac{m}2
%   %     }{
%   %       (n + 1 + \tfrac{m}2)((n+1) + (n + \tfrac{m}2) \log{(n + \tfrac{m}2)})
%   %     }
%   %   \biggr\rvert
%   = 0 < 1
%   \,, $$
% since
% $$
% \frac{\log{(x + a + 1)}}{x \log{(x + a)}}
%   \sim \frac{\log{(x+1)}}{(x-a) \log{x}}
%   \sim \frac{\log{(x+1)}}{x \log{x}}
%   \sim \frac{\tfrac1{(x+1)}}{1 + \log{x}}
%   \to 0
%   \,. $$
% A theorem from calculus states, that the formal series derivative (integral)
% coincides with the derivative (integral) of the function, corresponding to
% the power series (everywhere on the convergence region). And the convergence
% regions of derivative (integral) coincide with the region of the original
% power series.

Differentiating the series within $g_m$ yields
\begin{equation*}  \label{eq:g-m-deriv}
  \frac{d}{d x} g_m(x)
    % = e^{-x} \sum_{n\geq 1} \frac{x^{n-1}}{(n-1)!} \psi(n + \tfrac{m}2) - g_m(x)
    % = e^{-x} \sum_{n\geq 0} \frac{x^n}{n!} \psi(n + \tfrac{m}2 + 1) - g_m(x)
    % = e^{-x} \sum_{n\geq 0} \frac{x^n}{n!} (
    %     \psi(n + \tfrac{m}2 + 1) - \psi(n + \tfrac{m}2)
    % )
    % = e^{-x} \sum_{n\geq 0} \frac{x^n}{n!} \tfrac1{n + \tfrac{m}2}
    = e^{-x} \sum_{n\geq 0} \frac{x^n}{n!} (n + \tfrac{m}2)^{-1}
    = \cdots
    \,.
\end{equation*}
Since $
    e^t  = \sum_{n\geq 0} \tfrac{t^n}{n!}
$ on $\mathbb{R}$, for $\alpha\neq 0$ we have $
  \int_0^x t^{\alpha-1} dt
    = \tfrac{x^\alpha}{\alpha}
$ and
\begin{equation*}  \label{eq:g-m-deriv-int-1}
  \sum_{n\geq 0} \int_0^x \frac{t^{n+\alpha-1}}{n!} dt
    = \int_0^x \sum_{n\geq 0} \frac{t^{n+\alpha-1}}{n!} dt
    = \int_0^x t^{\alpha-1} e^t dt
    \,,
  % the MCT ([0, x], \mathcal{B}([0, x]), dx)
\end{equation*}
by the monotone convergence theorem, whence
\begin{equation}  \label{eq:g-m-deriv-int-2}
  \cdots
    = x^{-\tfrac{m}2} e^{-x} \sum_{n\geq 0}
        \frac{x^{n + \tfrac{m}2}}{n!} (n + \tfrac{m}2)^{-1}
    % = x^{-\tfrac{m}2} e^{-x} \sum_{n\geq 0}
    %     \frac1{n!} \int_0^x t^{n + \tfrac{m}2 - 1} dt
    % = x^{-\tfrac{m}2} e^{-x}
    %     \int_0^x t^{\tfrac{m}2 - 1} \sum_{n\geq 0} \frac{t^n}{n!} dt
    = x^{-\tfrac{m}2} e^{-x}
        \int_0^x t^{\tfrac{m}2 - 1} e^t dt
    = \cdots
    \,.
\end{equation}
The substitution $u^2 = t$ on $[0, \infty]$ with $2u du = dt$ yields
$$
\frac{d}{d x} g_m(x)
  = x^{-\tfrac{m}2} e^{-x}
      \int_0^x t^{\tfrac{m}2 - 1} e^t dt
  = 2 x^{-\tfrac{m}2} e^{-x}
      \int_0^{\sqrt{x}} u^{m - 1} e^{u^2} du
  \,. $$
In particular, for $m=1$ we get
\begin{equation}  \label{eq:g-m-series-one}
  g_1(x)
    = - \gamma - 2 \log{2}
    + e^{-x} \sum_{n\geq 0} \frac{x^n}{n!} \sum_{p=1}^n \frac2{2p - 1}
    \,,
\end{equation}
and the derivative \eqref{eq:g-m-deriv} can be computed thus
\begin{equation}  \label{eq:g-m-deriv-one}
  \frac{d}{d x} g_1(x)
    = 2 \tfrac{F(\sqrt{x})}{\sqrt{x}}
    \,,
\end{equation}
using $
  F \colon \mathbb{R} \to \mathbb{R}
  \colon x \mapsto e^{-x^2} \int_0^x e^{u^2} du
$ (Dawson's integral) a special function.

The derivative of \eqref{eq:ard-kl-div-real} with respect to $\alpha$ is
$$
\tfrac{d}{d \alpha} g_1(\tfrac1{2\alpha})
  = -\tfrac1{2 \alpha^2} g_1'(\tfrac1{2\alpha})
  % = -\tfrac1{2 \alpha^2} 2 \tfrac{F(\sqrt{\tfrac1{2\alpha}})}{\sqrt{\tfrac1{2\alpha}}}
  % = -\tfrac1{2 \alpha^2} 2 \tfrac{F(\tfrac1{\sqrt{2\alpha}})}{\tfrac1{\sqrt{2\alpha}}}
  = -\tfrac1{\alpha} \sqrt{\tfrac2{\alpha}} F(\tfrac1{\sqrt{2\alpha}})
  \,. $$
However, since $\alpha$ is nonnegative, it is typically parameterized via its logarithm
and computed when needed. Thus the gradient w.r.t $\log \alpha$ is
\begin{equation}  \label{eq:real-kl-div-deriv-log}
  \frac{d}{d\log \alpha}
    \tfrac12 \mathbb{E}_{z \sim \mathcal{N}(0,1)}
    \log \lvert z + \tfrac1{\sqrt{\alpha}} \rvert^2
    = \frac12 \frac{d\alpha}{d\log \alpha}
      \frac{d}{d\alpha} \bigl(\mathbb{E}\cdots \bigr)
    % = - \frac\alpha2 \tfrac1{\alpha}
    %   \sqrt{\tfrac2{\alpha}} F(\tfrac1{\sqrt{2\alpha}})
    = - \tfrac1{\sqrt{2\alpha}} F(\tfrac1{\sqrt{2\alpha}})
    \,.
\end{equation}
